\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-08 pldi/draft.tex
More intro

The \unix shell is an environment---often interactive---for composing scripts written in a plethora of programming languages.
This language agnosticism, coupled with \unix's toolbox philosophy~\cite{mcIlroymemo}, makes the shell the primary choice for specifying succinct and simple pipelines that target data processing, system orchestration, and other automation tasks.
% Pipe-centric shell commands are parallel (and to some extent, functional) programs that can natively enjoy significant parallelism on a multicore machine.
Unfortunately, parallelizing such pipelines requires significant effort and is shared unevenly between two different groups (or roles) of a programmer. % it's more like personas or "wearing hats"

The first group is \emph{command developers} responsible for developing individual commands such as \ttt{sort}, \ttt{uniq}, and \ttt{jq}.
These developers usually work in a single---often low-level---programming language and manually incorporate parallelism through means available in their chosen language, whenever this is possible. %  by the problem and language.
% the language and abstractions they are using focused at a particular task 
% also: target a small problem but aim at high reusability 
As they have no visibility into the command's uses, they expose a plethora of ad-hoc command-specific flags such as \ttt{-t}~\cite{}, \ttt{--parallel}~\cite{}, \ttt{-p}, \ttt{CPUs}, and \ttt{NUM\_THREADS}.

The second group is \emph{command users}, composing scripts out of individual commands.
% Unfortunately, users do not operate in a single language
As their programs combine a variety of commands written in many languages, they are left with only a few manual options for achieving parallelism:
% use the composition abstractions of the \unix shell, 
% and are often prototypical in nature---focusing on quickly prototyping a solution for the task at hand rather building an optimal solution to a recurring pattern.
% short lived, changing quickly
either coarse-grained tools such as \eg GNU \ttt{parallel}~\cite{Tange2011a}, \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm}, or manual parallelism POSIX shell primitives and 
% (ii) ad-hoc flags provided by command developers and mentioned above.

Their last resort is 
and writing the program from scratch
Manual rewriting is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality---but also foregoes all the productivity benefits of the original script.
% The fact that a programmer often belongs to both groups at the same time---as composed scripts can themselves form a command used in other scrips---compounds the challenge.
Often, these users are domain experts lacking a deep understanding of parallel systems---in desperate need of automation when it comes to parallelization.

This paper presents a solution targeting the second class of programmers, command users, 
\sys can automate

A system that is aware of the \unix philosophy, its key abstractions, and offers automated POSIX-aware parallelization as long as 
These lightweight annotations can be expressed by developers 
can be expressed 
This is paired with parallelizability study that informed the design of the system and its annotation interface, and led to the parallelizability semantics for POSIX and GNU coreutils---a large class of commands that serves as the standard library.

@@

To address this challenge, we develop \sys, a system for transforming POSIX shell scripts to their parallel equivalents.
\sys packs several ideas around the insight that a large subset of the Shell language can be represented as a dataflow graph, enabling for parallelization optimizations.
First, a careful study of the parallelizability properties of shell primitives and commands.
\sys addresses the challenge by offering command developers a unified set of abstractions for describing the parallelizability characteristics of their commands.
Using lightweight annotations akin to type signatures, \sys can automate the parallelization of entire POSIX scripts without any input from their user.
hand-in-hand with the parallelizability study:
  the study informed the annotation language and was used to annotate the entire
  for all 
---a set of annotations that \sys already offers for 
for the standard library
  akin to the standard library
  the annotation language is informed by the study, but also the study is 
XX to a type signature, these annotations describe

Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
(one sentence for the model)
(one sentence for the optimizations)
The result of the compilation is still a POSIX-compliant script that manages parallelism explicitly;
  POSIX compliance means 
it does not depend on a managed runtime

Third, \sys packs a runtime component 
A small library of highly-optimized data combiners allows 
This is paired with a
it comes with a few runtime primitives for eager data splitting and merging that address several practical issues ranging from deadlocks, performance bottlenecks;
  these primitives were developed from uncovering several pathological situations, on many of which we report.

We perform an extensive evaluation of \sys using 
  (i) a series of benchmarks, ranging from classic scripts found in the \unix literature to modern data-processing pipelines;
  (ii) three large and complex use cases, ranging from temperature analysis to web crawling and indexing; and
  (iii) a series of micro-benchmarks inspired by (and comparing with) possible alternatives.
The results demonstrate significant advantages in correctness, manual effort, and performance---with \sys's transformations offering an order-of-magnitude performance improvements aided further by \sys's optimized runtime components.
Improvements on the performance of individual commands depend on several factors such as the script's execution time, the pipeline depth and the parallelizability class under which commands fall.

>>>>

The \unix shell is an environment---often interactive---for composing scripts written in a plethora of programming languages.
This language-agnosticism, coupled with \unix's toolbox philosophy~\cite{mcilroy1978unix}, makes the shell the primary choice for specifying succinct and simple pipelines that target data processing, system orchestration, and other automation tasks.
% Pipe-centric shell commands are parallel (and to some extent, functional) programs that can natively enjoy significant parallelism on a multicore machine.
Unfortunately, parallelizing such pipelines requires significant effort shared between two different programmer groups. % it's more like roles personas or "wearing hats"

The first group is \emph{command developers}, responsible for developing individual commands such as \ttt{sort}, \ttt{uniq}, and \ttt{jq}.
These developers usually work in a single programming language, leveraging its abstractions to provide parallelism whenever possible.
As they have no visibility into the command's uses, they expose a plethora of ad-hoc command-specific flags such as \ttt{-t}, \ttt{-}\ttt{-parallel}, \ttt{-p}, and \ttt{-j} % , \ttt{CPUs}, and \ttt{NUM\_THREADS}~\cite{}.
\cite{pasetto2011comparative, mcilroy1993engineering, stallman1991gnu}.

The second group is \emph{command users}, who use POSIX shell constructs to combine multiple such commands from many languages into their scripts and are thus left with only a few options for incorporating parallelism (see~\S\ref{related}).
% If versed into parallelism, 
% Users versed in computing can apply 
% these tools are command-unaware thus at risk of breaking program semantics,
% require manual changes to the code to make parallelism explicit,
% at times leverage command flags like the ones mentioned earlier---but each command 
% oblivious? agnostic?  
One option is manual tools such GNU \ttt{parallel}~\cite{Tange2011a}, \ttt{tsc}~\cite{}, \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm};
  these options are either command-unaware thus at risk of breaking program semantics, or too coarse-grained thus applying at the level of entire scripts rather than individual components.
% Tools will attempt to parallelize anything 
% use the composition abstractions of the \unix shell, 
% and are often prototypical in nature---focusing on quickly prototyping a solution for the task at hand rather building an optimal solution to a recurring pattern.
% short lived, changing quickly
% (ii) ad-hoc flags provided by command developers and mentioned above.
Another option is the shell's concurrency primitives (\eg \ttt{\&}, \ttt{wait}, \ttt{for}) % for explicit parallelism, 
 that come at a cost of manual effort to carefully split inputs, manually rewrite scripts, and explicitly orchestrate execution---an expensive and error-prone process.
% re-writing of the entire script from scratch; unfortunately, this process is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality---but also foregoes all the productivity benefits of the original script.
% The fact that a programmer often belongs to both groups at the same time---as composed scripts can themselves form a command used in other scrips---compounds the challenge.
And all these options assume a good understanding of parallelism;
  users whose domain of expertise falls outside computing---routinely writing shell scripts to accomplish their tasks---are left without options.% automation when it comes to parallelization.

To address this challenge, we develop \sys, a shell transforming POSIX shell scripts to their parallel equivalents.
% \sys offers both classes of users in the second group such much-needed automation.
\sys benefits both programmer groups mentioned earlier, with a particular emphasis on users.
Command developers are provided a set of abstractions, akin to lightweight type annotations, for expressing the parallelizability properties of their commands:
  rather than providing a command's full observable behavior, these annotations focus primarily on their interaction with state.
Command users are provided full automation:
  \sys extracts latent parallelism from scripts as long as it understands their parallelizability properties.
% benefiting scripting experts and non-experts alike by offering automated POSIX-aware parallelization of shell-scripts as long as command developers
Its transformations are conservative, in that they do not attempt to parallelize fragments that lack sufficient information---\ie at worst, \sys will not improve performance rather than risking breakage.

To address cold-start issues, \sys comes with a library of parallelizability semantics for POSIX and GNU commands.
These large classes of commands serve as the shell's standard library and are used pervasively. % in scripts found out in the wild.
The study that led to the characterization of these commands also informed \sys's annotation language and transformation infrastructure.\footnote{
  In fact, the scope of our study is somewhat broader as it also incorporates frequently used commands that fall outside GNU Coreutils and POSIX such as \ttt{curl} and \ttt{pandoc}.
}

The parallelization study and transformations are paired with \sys's runtime component aware of the \unix philosophy and its key abstractions.
% \sys's runtime  component additional packs
% Aside from these two components, \sys packs a runtime component that serves dual purpose:
This component packs a small library of highly-optimized data combiners as well as high-performance primitives for eager data splitting and merging.
It addresses several practical challenges such as task parallelism, deadlocks, and performance bottlenecks, and was developed by uncovering several pathological situations, on many of which we report.

@@

% \sys packs several ideas around the insight that a large subset of the Shell language can be represented as a dataflow graph, enabling for parallelization optimizations.
% First, a careful study of the parallelizability properties of shell primitives and commands.
% \sys addresses the challenge by offering command developers a unified set of abstractions for describing the parallelizability characteristics of their commands.
% Using lightweight annotations akin to type signatures, \sys can automate the parallelization of entire POSIX scripts without any input from their user.
% hand-in-hand with the parallelizability study:
%   the study informed the annotation language and was used to annotate the entire
%   for all 
% ---a set of annotations that \sys already offers for 
% for the standard library
%   akin to the standard library
%   the annotation language is informed by the study, but also the study is 
% XX to a type signature, these annotations describe 
% 
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
% (one sentence for the model)
% (one sentence for the optimizations)
% The result of the compilation is still a POSIX-compliant script that manages parallelism explicitly;
%   POSIX compliance means 
% it does not depend on a managed runtime 
% 
% Third, \sys packs a runtime component 
% A small library of highly-optimized data combiners allows 
% This is paired with a
% it comes with a few runtime primitives for eager data splitting and merging that address several practical issues ranging from deadlocks, performance bottlenecks;
%   these primitives were developed from uncovering several pathological situations, on many of which we report.

We evaluate \sys using
  (i) a series of benchmarks, ranging from classic one-liners found in the \unix literature to modern data-processing pipelines;
  (ii) three large and complex use cases, covering temperature analysis, web crawling and indexing, and container orchestration; and
  (iii) a series of targeted micro-benchmarks inspired by (and comparing with) possible alternatives.
The results demonstrate significant advantages in correctness, manual effort, and performance---with \sys's transformations offering an order-of-magnitude performance improvements aided further by \sys's optimized runtime components.
Naturally, improvements on individual commands depend on several factors such as the script's execution time, the pipeline depth, and the parallelizability class under which a command falls.
